---
title: "Lab 1"
author: "Bushong Boys (Group 17)"
date: "1/20/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
```

## Exercise 1


Comparing disp to hwy:

In the scatterplot below, 38 models of cars engine sizes are compared to fuel efficiency on the highway. As you can see, there is an inverse relationship between these two variables. This makes logical sense since as the engine size of a car goes up, so does it actual size. Generally the larger a car is, the worse its fuel efficiency. 

```{r, echo = FALSE}
# Assign mpg to a variable  
data1 = mpg

# Create a plot comparing disp and hwy
v1 <- ggplot(data = data1, mapping = aes(x = displ, y = hwy)) + 
  geom_point()

plot(v1)
```


Comparing class to drv: 

The plot below compares a car's class to its type of drive. As seen by the graph, there is no inherent relationship between these two things since they are features of a car. There are plenty of SUVs with both front, rear, and four-wheel drive. Therefore, this a very un-useful visualization.

```{r, echo = FALSE}
# Create a plot comparing class and drv
v2 <- ggplot(data = data1, mapping = aes(x = class, y = drv)) + 
  geom_point()

plot(v2)
```

Back to disp and hwy: 

You may notice from our previous scatterplot comparing engine size and fuel efficiency that there is a group of cars with large engine sizes but have a higher fuel efficiency than a linear model might have predicted. One of our hypotheses was that the type of car was influencing this characteristic. Once adding the class of car as another level in the form of color, it was revealed that all of the outlier points were two seaters cars. After putting thought into this fact, it makes sense since they are smaller cars, but have larger engines for sporting capabilities.

```{r, echo = FALSE}
#Exersize 1b - add a color level of class to v1 
v3 <- ggplot(data = data1, mapping = aes(x = displ, y = hwy, color = class)) + 
  geom_point()

plot(v3)
```

## Exercise 2 

### Recommendation 1 - Timing the Campaign

Description of the Graph: 

Below, you will see a bar chart illustrating the month clients were called for an advertising campaign and the success rate of subscription from it. In addition, the black number above each bar is the number of clients reached during that month. For example, in October, 80 people were called to advertise the banks long-term savings account. 

Advice: 

Based on the results of this visualization, there are a few things we should observe. It would be naive to instantly conclude that the marketing campaign should take place during October, December, March, and September because this is when the success rate is the highest. The reasoning is these months have relatively small sample sizes and the actual mean might deviate far from the sample success rate. With that in mind, we recommend looking at a period of 3-4 months and there are two options based tolerance of risk. The first would be to choose the months October - December. This choice has the highest subscription success rate; however, three of these months have sample sizes under 80 which could yield in results that do not replicate this data. The more conservative approach would be to run the campaign February-April. This option still has high subscription success rates, but the greater sample sizes in February and April ensure a greater chance of success rates similar to these numbers.
 


```{r pressure, echo=FALSE}
# Load Bank Data
bankData <- read.csv("bank.csv")

# Subset data to only include the month and subscription success
month_y <- as_tibble(bankData[,c("month","y")])


# Creates a new data frame giving for each month the number of calls per month, number of subscriptions, number of rejections, and the success rate  
successRateFrame <- month_y %>%
  group_by(month) %>% 
  summarize(Count = n(), Y = sum(y == "yes"), N = sum(y == "no"), Success_Rate = sum(y == "yes")/n()) 

# Creates a bar graph that maps the success rate for each month. In addition, it includes the number of people who were contacted above 
successRatePlot <- ggplot(data = successRateFrame, mapping = aes(x= reorder(month, -Success_Rate) , y = Success_Rate)) + 
  geom_bar(stat = "identity", color = "purple", fill = "purple") + 
  labs(x = "Month", y = "Subscription Success Rate",title = "Subscription Success Rate by Month") + 
  theme(plot.title = element_text(hjust = .5, size = 18)) + 
  geom_text(
    aes(label=round(Count)), 
    size = 5, fontface = 2, color = 'black', 
    hjust=0.5, vjust=-1) + 
  ylim(0,.5)

plot(successRatePlot)
```

The next graph compares job type to Subscription success rate in an attempt to discover which people should be marketed to most heavily. 


```{r, echo=FALSE}
# Subset data to only include the job and subscription success
job_y <- as_tibble(bankData[,c("job","y")])


# Creates a new data frame giving for each job a client might have and counts the number of people in the data set who have a job category, number of subscriptions, number of rejections, and the success rate  
successRateFrame <- job_y %>%
  group_by(job) %>% 
  summarize(Count = n(), Y = sum(y == "yes"), N = sum(y == "no"), Success_Rate = sum(y == "yes")/n()) 

# Creates a bar graph that maps the success rate for each job category In addition, it includes the number of people who were contacted above 
jobSuccessRatePlot <- ggplot(data = successRateFrame, mapping = aes(x= reorder(job, -Success_Rate) , y = Success_Rate)) + 
  geom_bar(stat = "identity", color = "red", fill = "red") + 
  labs(x = "Job Category", y = "Subscription Success Rate",title = "Subscription Success Rate by Job Category") + 
  theme(plot.title = element_text(hjust = .5, size = 18), axis.text.x = element_text(angle=90, hjust=1)) + 
  geom_text(
    aes(label=round(Count)), 
    size = 5, fontface = 2, color = 'black', 
    hjust=0.5, vjust=-1) + 
  ylim(0,.5)

plot(jobSuccessRatePlot)
```

As you can see from the above chart, retirees and students are both excellent markets for long-term savings account campaigns. Generally, it appears white-collar workers are also more responsive as compared to blue-collar workers. We recommend targeting retirees, students, and white-collar workers most heavily to maximize the impact of the marketing campaign.